Estimating the number of clusters using Cross Validation

### 2015-08-15 cvclust.tex
add one more step in proof

\item the support of $P^Y$ contains at least two points.
\end{enumerate}
Then for any fixed $k > 1$, almost surely eventually  $\CV(k) > \CV(1)$.
\end{theorem}
\begin{proof}
Set $\muY = \E(Y)$ and $\SigmaY = \cov(Y)$.  The strong law of large
numbers implies that
\(
  \CV(1) \toas \int \|y - \mu^Y\|^2 P(dy) = \tr(\SigmaY).
\)
If $k > 1$, then Theorem~\ref{thm:kmeans-consistency} and
Lemma~\ref{lem:kmeans-separation} imply that there exists an $r > 0$ such that
almost surely eventually at most one cluster center in
$\{ \bmuY_1, \dotsc, \bmuY_k \}$ is within distance $r$ of $\muY$.  Without
loss of generality suppose that $\bmuY_1$ is the closest cluster center to
$\muY$. [TODO: finish proof]
\end{proof}

>>>>

\item the support of $P^Y$ contains at least two points.
\end{enumerate}
Then for any fixed $k > 1$, almost surely eventually  $\CV(k) > \CV(1)$.
\end{theorem}
\begin{proof}
Set $\muY = \E(Y)$ and $\SigmaY = \cov(Y)$.  The strong law of large
numbers implies that
\(
  \CV(1) \toas \int \|y - \mu^Y\|^2 P(dy) = \tr(\SigmaY).
\)

Now suppose that $k > 1$.  Let $\bmuY_1, \dotsc, \bmuY_k$ be the cluster
centers gotten from applying $k$-means to $Y_{n,1}, \dotsc, Y_{n,n}$.  Since
the support of $P^Y$ has at least two points,
Lemma~\ref{lem:kmeans-existence}, implies that almost surely eventually at
least two of the clusters will be nonempty and have distinct centers.  Without
loss of generality, assume that $k$ is less than or equal to the cardinality
of the support of $P^Y$.  Theorem~\ref{thm:kmeans-consistency} and
Lemma~\ref{lem:kmeans-separation} imply that there exists an $r > 0$ such that
almost surely eventually at most one cluster center is within distance $r$ of
$\muY$.  Without loss of generality suppose that $\bmuY_1$ is the closest
cluster center to $\muY$.

Let $\bmuX_1, \dotsc, \bmuX_k$ be the corresponding sample means computed from
$X_{n,1}, \dotsc, X_{n,n}$, with the $i$th observation assigned to cluster
$\hGY_{n,i} = \argmin_g \| Y_{n,i} - \bmuY_{n,i} \|^2.$  Consider the
classification rule $\hGY(X)$, which assigns $X$ to the class $g$ minimizing
$\|X - \bmuX_g\|^2$, breaking ties uniformly at random between classes of
equal distance.

[TODO: finish]

\end{proof}
